### 2013-01-31 htdocs/electrokinetic.tex
Revamped documentation
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@1902 40876441-3496-0410-a613-a2fca35b2ae2

\documentclass[aps,pre,preprint,groupedaddress]{revtex4-1}
\bibliographystyle{apsrev4-1}

\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
\newcommand{\e}[1]{\cdot10^{#1}}
\let\endtitlepage\relax

\begin{document}

\title{Report HPC-Europa2 Transnational Access Programme}

\author{Oliver Henrich$^1$ and Kevin Stratford$^2$}

\affiliation{
$^1$Centre for Computational Science, University College London, United Kingdom\\
$^2$Edinburgh Parallel Computing Centre, University of Edinburgh, United Kingdom
}

\date{\today}
\begin{abstract}
This is the intermim report of our joint visit from 12 April - 11 May 2012 to Barcelona Supercomputing Center and Prof Ignacio Pagonabarraga, Departament de Fisica Fonamental, Universitat de Barcelona, Spain.
\end{abstract}

\maketitle

\section{Week 1 (including period 12-20 April)}

We have introduced an object to hold the scalar electric potential,
and the charge densities for a general number of species. This
includes specification of diffusivities, valencies and unit charge.
The corresponding parallel halo swaps and I/O routines have been
added and subjected to unit tests.

In discussions with Pagonabarraga, we decided to prioritise the
implemenation of the the Poisson solver (successive overrelaxation,
SOR) so that more simple validation problems can be attempted as
soon as possible. The SOR solver has been implemented and tested
against an exact matrix solver for a one-dimensional problem in
both serial and parallel. The SOR solver assumes a spatially
uniform permittivity. We will extend this to a spatially varying
permittivity at a later stage.

A start has been made on the finite difference solution of the
Nernst Planck equation (advection/diffusion equation for charged
species). This will be completed and tested next week. This puts
us slightly ahead of schedule cf the original work plan.

Basic tests of the SOR solver have been already performed.
We have identified a number of other suitable test scenarios to 
verify the implemention of the Nernst-Planck equation in 
combination with the SOR solver.
We decided to test the Gouy-Chapman theory
for electric double layers in front of a charged wall \cite{Lyklema},
the liquid junction potential emerging between two electrolytes
of slighty different concentration and diffusivity of the charged species \cite{Mafe},
electro-osmotic flow in a slit pore  
\cite{Capuani, Rotenberg} and Debye-H\"uckel theory for charged 
colloidal particles and small enough potentials \cite{Lyklema}.

\section{Week 2 (23-27 April)}

We completed the numerical implementation of the Nernst-Planck
Equation which is solved via a simple finite difference method.
The solution includes advection (although there is no fluid
fluid fluid present in the test problems). We have included
no normal flux boundary conditions at solid fluid interfaces
to ensure the conservation of charged species. The solver
has been tested in both serial and in parallel.

A first validation test is a flat surface carrying a surface charge $\sigma$
with counterions and symmetic electrolyte. This is a one-dimensional, 
electroneutral problem of a diffusive electric double layer which has 
an analytical solution \cite{Lyklema}. 
The approximation for low electrostatic 
potentials reads

@@

@@

\begin{figure}[h!t]
\includegraphics[width=0.495\textwidth]{test1.pdf}
\includegraphics[width=0.495\textwidth]{test2.pdf}
\caption{Gouy-Chapman theory for electric double layers: The plots compare the simulation results for the electric potential with the analytical solution for $\rho_{0,\pm}=1\e{-2}$ (left) and $\rho_{0,\pm}=1\e{-3}$ (right).} 
\label{fig1} 
\end{figure}

@@

\begin{figure}[h!t]
\includegraphics[width=0.495\textwidth]{test3.pdf}
\includegraphics[width=0.495\textwidth]{test3-rho.pdf}
\caption{Nonlinear regime: For larger surface charges the solution deviates from the low-potential approximation Eq. \ref{gouychapman} to the Gouy-Chapman theory. The righthand picture shows the charge distribution across the gap.} 
\label{fig2} 
\end{figure}

A number of minor bugs in the parallel implementation of the
SOR solver were resolved, and the unit tests updated appropriately.

Some effort has been used to create a comprehensive test suite
for the basic functionality involved in these routines. These
tests are valid both in serial and in parallel.

Work is now under way to couple the new Nernst-Planck and
Poisson routines with the main code. This includes work to
include charged colloids.

\section{Week 3 (30 April - 4 May)}

We have now coupled the Poisson solver and the Nernst Planck solver
to the main code and checked for correct operation. This has included
the run time intialisation of the electrokinetic-related quatities
from the input file. We have also integrated the parallel I/O for
the electrokinetic lattice quantities to the main code, again with
appropriate initialisation from input.

We proceeded with another test. The liquid junction potential 
is a charge separation process that 
occurs when electrolytes with slightly different concentrations
whose species have different diffusivities are brought into contact.
Charges from the regions of higher concentration diffuse   
into the parts with lower concentration. Due to the difference 
in diffusivity they migrate at different speeds, leaving parts of
the system charged. This leads to a build-up of a potential
which balances the diffusive flux.

@@

@@

@@

A complete discussion of the solution can be found in \cite{Mafe}. 
There, the upper limit of significant modes has been also estimated as $N_{max} = L/\pi l_D$.
Note the factor 2 difference between Eq. \ref{taud} and the corresponding expression in \cite{Mafe}.
\begin{figure}[h!t]
\includegraphics[width=0.495\textwidth]{test_lj_zoom1.pdf}
\includegraphics[width=0.495\textwidth]{test_lj_zoom3.pdf}
\caption{Time evolution of the liquid junction potential for $L_x=128$ (blue), $L_x=256$ (green) and $L_x=512$ (blue). The dashed black curves represent the approximate solution in the limit $l_D/L_x<<1$. Deviations can be seen which are due
to the finite size of our system and the approximate nature of the analytical solution.} 
\label{fig3} 
\end{figure}

@@

\begin{figure}[h!t]
\includegraphics[width=0.9\textwidth]{test_lj_decay1.pdf}
\caption{Rescaled plot of the decay: Times have been Time evolution of the liquid junction potential for $L_x=128$ (blue), $L_x=256$ (green) and $L_x=512$ (blue). The dashed black curves represent the approximate solution in the limit $l_D/L_x<<1$. Deviations can be seen which are due
to the finite size of our system and the approximate nature of the analytical solution.} 
\label{fig4} 
\end{figure}

@@

A reason for this might be that the analytical solution was gained
for an infinitely large system with constant charge concentrations,
vanishing currents at both ends and finite diffusive zone of size $L_x$.
In our situation the entire system is within the diffusive zone.
This may lead to smaller effective diffusivities or larger effective
system sizes.
Interestingly, all runs with solid walls at both ends resulted in 
oscillatory behaviour and an effective system size of $2L_x$.

While the scaling for the Nernst Planck part of
the calulcation are reasonable, it can be seen that the SOR routine
used to solve the Poisson equation is not really acceptable. This
was not unexpected. Some improvements might be possible, but the
algorithm is essentailly unsuitable for large systems. We will
investigate the use of an alternative (e.g., multigrid), but this
is beyond the scope of this HPC Europa visit.

In discussion with Pagonabarraga, we have also made some progress
in deciding how to modify a number of other parts of the code to
allow for the presence of the electric potential. These include
the free energy calculation, which can be generalised to compute
both electric contributions and coupling between potential and
compositional order parameter. A flexible solution may require
some refactoring.

We have also implemented a fixed external electric field in the
Nernst Planck equation. This will be tested during the last week
of the visit.

\begin{figure}[h!t]
\includegraphics[width=0.9\textwidth]{scaling.pdf}
\caption{Scaling of a system of $128^3$ lattice sites on up to 512 MPI tasks. 
While the scaling for the Nernst Planck part of the calulcation are reasonable, 
it can be seen that the SOR routine used to solve the Poisson equation deviates
strongly from ideal behaviour.} 
\label{fig5} 
\end{figure}

\section{Week 4 (7 May - 11 May)}

@@

\begin{figure}[h!t]
\includegraphics[width=0.9\textwidth]{test_eo.pdf}
\caption{Steady state flow profile across a slit of width $L_x=63$ for an applied field of magnitude $e \beta E L_x=1.89$ for two different charge densities $\sigma=0.003125$ (red) and $\sigma=0.0125$ (blue), Bjerrum length $l_B=0.7234$, viscosity $\eta=0.1$ and unit charge $e=1$. The dashed black lines correspond to theoretical prediction according to Eqs. \ref{vy} and \ref{vhat}.} 
\label{fig6} 
\end{figure}

Finally, we have tested the implementation for a single fixed
colloid and compared the result with Debye-H\"uckel theory. A result
is shown in Fig.~\ref{fig7}.
We used the following parametrisation:

@@

\begin{figure}[h!t]
\includegraphics[angle=-90,width=0.9\textwidth]{test_dh1.pdf}\\
\includegraphics[angle=-90,width=0.9\textwidth]{test_dh2.pdf}
\caption{Debye-H\"uckel theory for positively charged colloid: the picture
shows a cut through the center of a single colloid in a peridoic system.
The potential is shown in red, the Stern potential in blue, and the negative 
charge density in green. The colloid is in the center.}
\label{fig7}
\end{figure}

\clearpage
\begin{thebibliography}{99}

\bibitem{Lyklema} J. Lyklema {\em Fundamentals of Interface and Colloid Science} Academic Press (1995).
\bibitem{Mafe} S. Maf\'e, J.A. Manzanares, J. Pellicer, {\em J. Electroanal. Chem.} {\bf 241}, 57-77 (1988).
\bibitem{Capuani} F. Capuani, I. Pagonabarraga, D. Frenkel, {\em J. Chem. Phys.} {\bf 121}, 973-986 (2004).
\bibitem{Rotenberg} B. Rotenberg, I. Pagonabarraga, D. Frenkel, {\em Farad. Discuss.} {\bf 144}, 223-243 (2010).

\end{thebibliography}

\end{document}

>>>>

\section{Electrokinetics}

\subsection{Sress tensor in the electrosymmetic model}

\subsubsection{Definitions}

The total free energy functional is a function of the composition $\phi$, the density of the ionic species $\rho_\alpha$,

\beq
{\cal F}[\phi,\{\rho_\alpha\}] = \int d{\bf r} \;F\left(\phi({\bf r}),\{\rho_\alpha({\bf r})\}\right). 
\eeq

The free energy density can be factorised into a contribution from the composition and an ionic contribution according to

\beqa
F&=&F^{mix} +F^{ion}\\
F^{mix}&=& \frac{1}{2} \,A \,\phi^2 + \frac{1}{4} \,B \,\phi^4+ \frac{1}{2} \kappa(\bm{\nabla}\phi)^2\\
F^{ion,ex}&=& \sum_{\alpha=\pm} \rho_\alpha \left(V^{solv}_\alpha-\mu_\alpha +\frac{z_\alpha e}{2}\Psi \right).
\eeqa

In the ionic contribution we have omitted the ideal gas contribution of the ions.
The solvation energy is given by

\beqa
V^{solv}_\alpha({\bf r})=\Delta\mu_\alpha\frac{1+\phi({\bf r})}{2}
\eeqa

It is convenient to define the total free energy without electric field $F_0$, i.e. with vanishing electric potential:

\beqa
F_0&=&F^{mix} +F^{ion,ex}|_{\Psi=0}
\eeqa

The Poisson equation reads
\beqa
\bm{\nabla}\left(\varepsilon({\bf r})\bm{\nabla}\Psi({\bf r})\right)=-\sum_{\alpha=\pm} e \,z_\alpha \,\rho_\alpha ({\bf r})
\eeqa

whereas the electric field is given by

\beq
{\bm E}=-{\bm \nabla}\Psi.
\eeq

The permittivity depends on the composition $\phi$ via

\beqa
\varepsilon({\bf r})=\bar{\varepsilon} \,(1-\gamma\,\phi({\bf r}))
\eeqa

The electric displacement field

\beq
{\bm D}({\bf r})=\varepsilon({\bf r}){\bm E}({\bf r})
\eeq

The 'co-energy' density

\beq
\tilde{F}=F_0-\frac{\varepsilon}{2}\bm{E}^2
\eeq

has the meaning of a Legendre transform of the free energy density $F_0$.

\subsubsection{Electromechanical stress tensor}

The electromechanical stress tensor is generally defined as \cite{Landau-ED, Melcher}.

\beqa
\sigma_{i j}=\varepsilon E_i E_j + \delta_{i j}\left( \tilde{F} - \phi \frac{\delta \tilde{\cal F}}{\delta\phi} - \sum_{\alpha=\pm} \rho_\alpha \frac{\delta \tilde{\cal F}}{\delta\rho_\alpha}\right).
\eeqa

The first functional derivative yields

\beqa
\phi \frac{\delta\tilde{\cal F}}{\delta\phi}&=&\phi\left(\frac{\delta {\cal F}_0}{\delta\phi}-\frac{{\bm E}^2}{2} \frac{\partial \varepsilon}{\partial\phi}\right)\\
&=&\phi\left(\frac{\partial F_0}{\partial\phi}-{\bm \nabla}\left(\frac{\partial F_0}{\partial {\bm \nabla}\phi}\right)-\frac{{\bm E}^2}{2} \frac{\partial \varepsilon}{\partial\phi}\right)\\
&=&A\,\phi^2+B\,\phi^4-\kappa\phi({\bm \nabla}^2\phi)+\frac{\phi}{2}\sum_{\alpha=\pm} \rho_\alpha \Delta\mu_\alpha+\phi\,\frac{\bar{\varepsilon}\gamma}{2}{\bm E}^2\\
&=&A\,\phi^2+B\,\phi^4-\kappa\phi({\bm \nabla}^2\phi)+\frac{\phi}{2}\sum_{\alpha=\pm} \rho_\alpha \Delta\mu_\alpha-\frac{\varepsilon-\bar{\varepsilon}}{2}{\bm E}^2,
\eeqa

where we used the above dependence of the permittivity on the composition.
The second functional derivative is given by

\beqa
\rho_\alpha \frac{\delta \tilde{\cal F}}{\delta\rho_\alpha}&=&\rho_\alpha\left(\frac{\delta {\cal F}_0}{\delta\rho_\alpha}\right)\\
&=&\rho_\alpha(V^{solv}_\alpha-\mu_\alpha)
\eeqa

The co-energy density and the functional derivatives with respect to composition $\phi$ and ionic densities $\rho_\alpha$ add up to

\beqa
\tilde{F} - \phi \frac{\delta \tilde{F}}{\delta\phi} - \sum_{\alpha=\pm} \rho_\alpha \frac{\delta \tilde{F}}{\delta\rho_\alpha}=&&\nonumber\\
&&\hspace*{-4cm}=-\frac{\varepsilon}{2}\bm{E}^2+\frac{A}{2} \phi^2 + \frac{B}{4} \phi^4+ \frac{\kappa}{2} (\bm{\nabla}\phi)^2 \nonumber\\
&&\hspace*{-3cm}-A\phi^2-B\phi^4+ \kappa\phi({\bm \nabla}^2\phi) -\frac{\phi}{2}\sum_{\alpha=\pm} \rho_\alpha \Delta\mu_\alpha\nonumber\\
&&\hspace*{-3cm}-\phi\,\frac{\bar{\varepsilon}\gamma}{2}{\bm E}^2-\sum_{\alpha=\pm}\rho_\alpha(V^{solv}_\alpha-\mu_\alpha)\\
&&\hspace*{-4cm}=-\frac{\varepsilon}{2}\bm{E}^2-\frac{A}{2}\phi^2 -\frac{3 B}{4} \phi^4+ \frac{\kappa}{2}(\bm{\nabla}\phi)^2 + \kappa\phi({\bm \nabla}^2\phi) \nonumber\\
&&\hspace*{-3cm}-\frac{\phi}{2}\sum_{\alpha=\pm} \rho_\alpha \Delta\mu_\alpha -\phi\,\frac{\bar{\varepsilon}\gamma}{2}{\bm E}^2
\eeqa

This yields

\beqa
\sigma_{ij}&=&\varepsilon E_i E_j - \delta_{i j}\left(\frac{\varepsilon}{2}\bm{E}^2+\frac{A}{2}\phi^2 +\frac{3\,B}{4}\phi^4 - \frac{\kappa}{2} (\bm{\nabla}\phi)^2\right.\nonumber\\
&&\left.- \kappa\phi({\bm \nabla}^2\phi) +\frac{\phi}{2}\sum_{\alpha=\pm} \rho_\alpha \Delta\mu_\alpha +\phi\,\frac{\bar{\varepsilon}\gamma}{2}{\bm E}^2\right).\nonumber\\
\eeqa

In order to satisfy the condition of mechanical equilibrium $\nabla_j \sigma_{ij}=0$ a symmetric contribution has to be added to the above stress tensor. The following lines try to elucidate this.\\

Taking the gradient of the stress tensor gives

\beqa
\nabla_j\sigma_{i j}&=& (\nabla_j \varepsilon E_j) E_i + \varepsilon E_j \nabla_j E_i + \nabla_i\left( \tilde{F} - \phi \frac{\delta \tilde{\cal F}}{\delta\phi} - \sum_{\alpha=\pm} \rho_\alpha \frac{\delta \tilde{\cal F}}{\delta\rho_\alpha}\right)\\
&=& \left({\bm \nabla}\cdot{\bm D}\right) E_i+ D_j \nabla_j E_i +\nabla_i F_0 - D_j \nabla_i E_j \nonumber\\
&&\hspace*{0.5cm} -\nabla_i\left(\phi \frac{\delta \tilde{\cal F}}{\delta\phi} +\sum_{\alpha=\pm} \rho_\alpha \frac{\delta \tilde{\cal F}}{\delta\rho_\alpha}\right)\\
&=& \left({\bm \nabla}\cdot{\bm D}\right) E_i + D_j \left(\nabla_j E_i-\nabla_i E_j\right) + \nabla_i F_0\nonumber\\
&&\hspace*{0.5cm} -\nabla_i\left(\phi \mu_\phi +\sum_{\alpha=\pm} \rho_\alpha \mu_{\rho_\alpha}\right)
\eeqa

According to Gauss' law the first term is

\beq
\left({\bm \nabla}\cdot{\bm D}\right)E_i = \rho_f E_i=\sum_{\alpha=\pm} z_\alpha e \rho_\alpha E_i,
\eeq

whereas $\rho_f$ is the free charge density.
Moreover, due to Faraday's law the curl in the second bracket vanishes.
This leads us to the following intermediate result:

\beqa
\nabla_j\sigma_{i j}&=& \sum_{\alpha=\pm}z_\alpha e \rho_\alpha E_i + \nabla_i F_0 - \nabla_i\left(\phi \mu_\phi +\sum_{\alpha=\pm} \rho_\alpha \mu_{\rho_\alpha}\right)\\
&=&\sum_{\alpha=\pm} z_\alpha e \rho_\alpha E_i + \frac{\partial F_0}{\partial \phi} \nabla_i \phi + \frac{\partial F_0}{\partial \nabla_j \phi} \nabla_j \nabla_i \phi +\sum_{\alpha=\pm} \left\{\frac{\partial F_0}{\partial \rho_\alpha}\right\} \nabla_i \rho_\alpha\nonumber\\
&&\hspace*{0.5cm} -\nabla_i\left(\phi \mu_\phi +\sum_{\alpha=\pm} \rho_\alpha \mu_{\rho_\alpha}\right)\\
&=& \sum_{\alpha=\pm} z_\alpha e \rho_\alpha E_i + \left\{\frac{\partial F_0}{\partial \phi} - \nabla_j \left(\frac{\partial F_0}{\partial \nabla_j \phi}\right)\right\} \nabla_i \phi +\nabla_j\left(\frac{\partial F_0}{\partial \nabla_j \phi} \nabla_i \phi\right)\nonumber\\
&&\hspace*{0.5cm} +\sum_{\alpha=\pm} \left\{\frac{\partial F_0}{\partial \rho_\alpha} \right\}\nabla_i \rho_\alpha-\nabla_i\left(\phi \mu_\phi +\sum_{\alpha=\pm} \rho_\alpha \mu_{\rho_\alpha}\right).
\eeqa

If we replace the terms in curly brackets the chemical potentials $\mu_\phi$ and $\mu_{\rho_\alpha}$ and use the fact that gradients of the chemical potential vanish in equilibrium, we end up with

\beqa
\nabla_j\sigma_{i j}&=& \sum_{\alpha=\pm}z_\alpha e \rho_\alpha E_i + \nabla_j\left(\frac{\partial F_0}{\partial \nabla_j \phi} \nabla_i \phi\right)
\eeqa

The first term is the external force on the charges, which is zero if we have no counterions and an equal amount of positve and negative ions. Therefore, we have to add a term

\beq
-\left(\frac{\partial F_0}{\partial \nabla_j \phi} \nabla_i \phi\right)
\eeq

to the stress tensor to fulfill the equilibrium condition. This is equivalent to similar derivations in \cite{Landau-EL}.\\

The total stress tensor reads

\beqa
\sigma_{ij}&=&\varepsilon E_i E_j - \kappa (\nabla_i\phi)(\nabla_j\phi)  - \delta_{i j}\left(\frac{\varepsilon}{2}\bm{E}^2+\frac{A}{2}\phi^2 +\frac{3\,B}{4}\phi^4\right.\nonumber\\
&&\left. - \frac{\kappa}{2} (\bm{\nabla}\phi)^2- \kappa\phi({\bm \nabla}^2\phi) +\frac{\phi}{2}\sum_{\alpha=\pm} \rho_\alpha \Delta\mu_\alpha +\phi\,\frac{\bar{\varepsilon}\gamma}{2}{\bm E}^2\right).
\eeqa

\subsection{Examples}

A number of regression tests 
verify the implemention of the Nernst-Planck equation in 
combination with the SOR solver. They can be found in

\begin{verbatim}
trunk/tests/regression
\end{verbatim}

The original tests we carried out during the first
implementation comprise the Gouy-Chapman theory
for electric double layers in front of a charged wall \cite{Lyklema},
the liquid junction potential emerging between two electrolytes
of slighty different concentration and diffusivity of the charged species \cite{Mafe},
electro-osmotic flow in a slit pore  \cite{Capuani, Rotenberg} 
and Debye-H\"uckel theory for charged 
colloidal particles and small enough potentials \cite{Lyklema}.
They are described in the following paragraphs.

\subsubsection{Gouy-Chapman}

This validation test is a flat surface carrying a surface charge $\sigma$
with counterions and symmetic electrolyte. This is a one-dimensional, 
electroneutral problem of a diffusive electric double layer which has 
an analytical solution \cite{Lyklema}. 
The approximation for low electrostatic 
potentials reads

\begin{figure}[htpb]
\includegraphics[width=0.495\textwidth]{./pics/test1.pdf}
\includegraphics[width=0.495\textwidth]{./pics/test2.pdf}
\caption{Gouy-Chapman theory for electric double layers: The plots compare the simulation results for the electric potential with the analytical solution for $\rho_{0,\pm}=1\e{-2}$ (left) and $\rho_{0,\pm}=1\e{-3}$ (right).} 
\label{fig1} 
\end{figure}

@@

\begin{figure}[htpb]
\includegraphics[width=0.495\textwidth]{./pics/test3.pdf}
\includegraphics[width=0.495\textwidth]{./pics/test3-rho.pdf}
\caption{Nonlinear regime: For larger surface charges the solution deviates from the low-potential approximation Eq. \ref{gouychapman} to the Gouy-Chapman theory. The righthand picture shows the charge distribution across the gap.} 
\label{fig2} 
\end{figure}

@@

@@

\subsubsection{Liquid-junction potential}

The liquid junction potential 
is a charge separation process that 
occurs when electrolytes with slightly different concentrations
whose species have different diffusivities are brought into contact.
Charges from the regions of higher concentration diffuse   
into the parts with lower concentration. Due to the difference 
in diffusivity they migrate at different speeds, leaving parts of
the system charged. This leads to a build-up of a potential
which balances the diffusive flux.

@@

\begin{figure}[h!t]
\includegraphics[width=0.495\textwidth]{./pics/test_lj_zoom1.pdf}
\includegraphics[width=0.495\textwidth]{./pics/test_lj_zoom3.pdf}
\caption{Time evolution of the liquid junction potential for $L_x=128$ (blue), $L_x=256$ (green) and $L_x=512$ (blue). The dashed black curves represent the approximate solution in the limit $l_D/L_x<<1$. Deviations can be seen which are presumably due
to the approximate nature of the analytical solution and the fact that we
have a different asymptotic flux close to the boundary - the theoretical 
analysis assumes no flux at a boundary which is infinitely far away from
the diffusive region.} 
\label{fig3} 
\end{figure}

@@

\begin{figure}[htpb]
\includegraphics[width=0.9\textwidth]{./pics/test_lj_decay1.pdf}
\caption{Rescaled plot of the decay: Times have been Time evolution of the liquid junction potential for $L_x=128$ (blue), $L_x=256$ (green) and $L_x=512$ (blue). The dashed black curves represent the approximate solution in the limit $l_D/L_x<<1$. Deviations can be seen which are due
to the approximate nature of the analytical solution and the diffusive fluxes close to the boundary.} 
\label{fig4} 
\end{figure}

@@

A complete discussion of the solution can be found in \cite{Mafe}. 
There, the upper limit of significant modes has been also estimated as $N_{max} = L/\pi l_D$.
Note the factor 2 difference between Eq. \ref{taud} and the corresponding expression in \cite{Mafe}.

@@

@@

A reason for this might be the approximate nature of the analytical solution 
and the fact that it was gained
for an infinitely large system with constant charge concentrations,
vanishing currents at both ends and finite diffusive zone of size $L_x$.
In our situation the entire system is within the diffusive zone.
This may lead to smaller effective diffusivities or larger effective
system sizes.
Interestingly, all runs with solid walls at both ends resulted in 
oscillatory behaviour and an effective system size of $2L_x$.

\subsubsection{Electroosmotic Flow}

@@

\begin{figure}[htpb]
\includegraphics[width=0.9\textwidth]{./pics/test_eo.pdf}
\caption{Steady state flow profile across a slit of width $L_x=63$ for an applied field of magnitude $e \beta E L_x=1.89$ for two different charge densities $\sigma=0.003125$ (red) and $\sigma=0.0125$ (blue), Bjerrum length $l_B=0.7234$, viscosity $\eta=0.1$ and unit charge $e=1$. The dashed black lines correspond to theoretical prediction according to Eqs. \ref{vy} and \ref{vhat}.} 
\label{fig6} 
\end{figure}

\subsubsection{Debye-H\"uckel Theory}

We have tested the implementation for a single fixed
colloid and compared the result with Debye-H\"uckel theory. A result
is shown in Fig.~\ref{fig7}.
We used the following parametrisation:

@@

\begin{figure}[htpb]
\includegraphics[angle=-90,width=0.9\textwidth]{./pics/test_dh1.pdf}\\
\includegraphics[angle=-90,width=0.9\textwidth]{./pics/test_dh2.pdf}
\caption{Debye-H\"uckel theory for positively charged colloid: the picture
shows a cut through the center of a single colloid in a peridoic system.
The potential is shown in red, the Stern potential in blue, and the negative 
charge density in green. The colloid is in the center.}
\label{fig7}
\end{figure}
\clearpage

\subsubsection{Scaling of the SOR}

While the scaling for the Nernst Planck part of
the calulcation are reasonable, it can be seen that the SOR routine
used to solve the Poisson equation is not really acceptable. This
was not unexpected. Some improvements might be possible, but the
algorithm is essentailly unsuitable for large systems.

\begin{figure}[htpb]
\includegraphics[width=0.9\textwidth]{./pics/scaling.pdf}
\caption{Scaling of a system of $128^3$ lattice sites on up to 512 MPI tasks. 
While the scaling for the Nernst Planck part of the calulcation are reasonable, 
it can be seen that the SOR routine used to solve the Poisson equation deviates
strongly from ideal behaviour.} 
\label{fig5} 
\end{figure}
\clearpage

\subsection{User input}

### 2013-01-31 htdocs/references.tex
Revamped documentation
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@1902 40876441-3496-0410-a613-a2fca35b2ae2

\begin{thebibliography}{99}

@@

@@

>>>>

\addcontentsline{toc}{section}{References}

\bibliographystyle{plain}
\begin{thebibliography}{99}

@@

\bibitem{fournier2005}
J.-B. Fournier and P. Galatola,
Modeling planar degenerate wetting and anchoring in nematic liquid
crystals,
\textit{Europhys. Lett.}, \textbf{72} 403--409 (2005).

@@

\bibitem{Lyklema} J. Lyklema {\em Fundamentals of Interface and Colloid Science} Academic Press     (1995).
\bibitem{Mafe} S. Maf\'e, J.A. Manzanares, J. Pellicer, {\textit J. Electroanal. Chem.} {\textbf 241}, 5    7-77 (1988).
\bibitem{Capuani} F. Capuani, I. Pagonabarraga, D. Frenkel, {\textit J. Chem. Phys.} {\textbf 121}, 973-    986 (2004).
\bibitem{Rotenberg} B. Rotenberg, I. Pagonabarraga, D. Frenkel, {\textit Farad. Discuss.} {\textbf 144},     223-243 (2010).
\bibitem{Landau-ED} L.D. Landau, E.M. Lifshitz, {\textit Electrodynamics of Continuous Media}, \S 15    , 2nd ed., Pergamon Press, Oxford, UK (1984).
\bibitem{Melcher} J.R. Melcher, {\textit Continuum Electromechanics}, \S 3.10, MIT Press, Cambridge,     MA, USA (1981).\\
downloadable from:\\
\url{http://ocw.mit.edu/ans7870/resources/melcher/resized/cem_811.pdf}
\bibitem{Landau-EL} L.D. Landau, E.M. Lifshitz, {\textit Theory of Elasticity}, \S 3 \& \S 16, 3rd e    d., Butterworth-Heinemann, Oxford, UK (1986).
